---
title: "Filter_Join"
output: html_notebook
---

https://rstudio-conf-2020.github.io/r-for-excel/filter-join.html

filter() to conditionally subset our data by rows, and
*_join() functions to merge data frames together
And we’ll make a nicely formatted HTML table with kable() and kableExtra
The combination of filter() and *_join() - to return rows satisfying a condition we specify, and merging data frames by like variables - is analogous to the useful VLOOKUP function in Excel.

Use filter() to subset data frames, returning rows that satisfy variable conditions
Use full_join(), left_join(), and inner_join() to merge data frames, with different endpoints in mind
Use filter() and *_join() as part of a wrangling sequence

```{r}
library(tidyverse)
library(readxl)
library(here) 
library(kableExtra)
```
```{r}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_excel(here("data", "kelp.xlsx"), sheet = "abur")
```

```{r}
View(fish)
```
```{r}
View(kelp_abur)
```
```{r}
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")

fish_garibaldi
```
```{r}
fish_mohk <- fish %>% 
  filter(site == "mohk")
fish_mohk
```
```{r}
fish_over50 <- fish %>% 
  filter(total_count >= 50)
fish_over50
```
```{r}
fish_3sp <- fish %>% 
  filter(common_name == "garibaldi" | 
           common_name == "blacksmith" | 
           common_name == "black surfperch")
fish_3sp
```
```{r}
fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))
```

```{r}
fish_gar_2016 <- fish %>% 
  filter(year == 2016 | common_name == "garibaldi")
```
```{r}
aque_2018 <- fish %>% 
  filter(year == 2018, site == "aque")
aque_2018

```

```{r}
aque_2018 <- fish %>% 
  filter(year == 2018 & site == "aque")
aque_2018
```

```{r}
# Written as sequential filter steps:

aque_2018 <- fish %>% 
  filter(year == 2018) %>% 
  filter(site == "aque")

aque_2018
```
```{r}
low_gb_wr <- fish %>% 
  filter(common_name %in% c("garibaldi", "rock wrasse"), 
         total_count <= 10)
```

stringr::str_detect() to filter by a partial pattern

```{r}
fish_bl <- fish %>% 
  filter(str_detect(common_name, pattern = "black"))
fish_bl
```
```{r}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern = "it"))

# blacksmITh and senorITa remain!
fish_it
```
```{r}
abur_kelp_fish <- kelp_abur %>% 
  full_join(fish, by = c("year", "site"))

abur_kelp_fish
```

```{r}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year","site"))
kelp_fish_left
```

```{r}

kelp_fish_injoin <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
kelp_fish_injoin
# kelp_fish_injoin
```
```{r}
my_fish_join <- fish %>% 
  filter(year == 2017, site == "abur") %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count / total_fronds)

my_fish_join
```
```{r}
kable(my_fish_join)

```

```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE)
```

