---
title: "Data reshaping"
output: html_notebook
---
Use tidyr::pivot_wider() and tidyr::pivot_longer() to reshape data frames
janitor::clean_names() to make column headers more manageable
tidyr::unite() and tidyr::separate() to merge or separate information from different columns
Detect or replace a string with stringr functions

https://rstudio-conf-2020.github.io/r-for-excel/tidying.html

```{r}
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```

tidyr::pivot_longer() to reshape from wider-to-longer format

```{r}
inverts <- read_excel(here("data", "inverts.xlsx"))
```

```{r}
View(inverts)
```
```{r}
summary(inverts)
```
```{r}
names(inverts)
```

```{r}
inverts_long <- pivot_longer(data = inverts, 
                                    cols = '2016':'2018',
                                    names_to = "year",
                                    values_to = "sp_count")
```

```{r}
View(inverts_long)
```
```{r}
summary(inverts_long)
```

```{r}
# Coerce "year" class to numeric: 

inverts_long <- inverts_long %>% 
  mutate(year = as.numeric(year))
```

```{r}
View(inverts_long)
```
```{r}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name, 
                     values_from = sp_count)
```

```{r}
View(inverts_wide)
```
```{r}
inverts_wide
```

The janitor package by Sam Firke is a great collection of functions for some quick data cleaning, like:

janitor::clean_names(): update column headers to a case of your choosing
janitor::get_dupes(): see all rows that are duplicates within variables you choose
janitor::remove_empty(): remove empty rows and/or columns
janitor::adorn_*(): jazz up tables

```{r}
inverts_wide <- inverts_wide %>% 
  clean_names()
```

```{r}
summary(inverts_wide)
```
tidyr::unite() and tidyr::separate() to combine or separate information in column(s)

```{r}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year", # What to name the new united column
               c(site, year), # The columns we'll unite (site, year)
               sep = "_") # How to separate the things we're uniting
inverts_unite
```
```{r}
inverts_moyr <- inverts_long %>% 
  unite(col = "mo_yr", # What to name the new united column
               c(month, year), # The columns we'll unite (site, year)
               sep = "/") 

inverts_moyr
```
```{r}
# Uniting more than 2 columns: 

inverts_triple_unite <- inverts_long %>% 
  tidyr::unite(col = "year_site_name",
               c(year, site, common_name),
               sep = "-") # Note: this is a dash

inverts_triple_unite
```
```{r}
inverts_sep <- inverts_unite %>% 
  tidyr::separate(site_year, into = c("my_site", "my_year"))
inverts_sep
```

stringr::str_replace() to replace a pattern

```{r}
ca_abbr <- inverts %>% 
  mutate(
    common_name = 
      str_replace(common_name, 
              pattern = "california", 
              replacement = "CA")
  )

ca_abbr
```

