---
title: "R for Excel Practice"
output: html_notebook
---
## Import basic library for data cleaning
here: library for pointing files at current project working directory
```{r}
library(tidyverse)
library(readxl)
library(here)
```
```{r}
ca_np <- read_csv(here("data", "ca_np.csv"))
```
```{r}
View(ca_np)
```
```{r}
names(ca_np)
```
```{r}
head(ca_np)
```
```{r}
summary(ca_np)
```
```{r}
library(readxl)
```
```{r}
##ci_np : national parks for Channel Islands
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```
```{r}
View(ci_np)
```

ggplot(data = df_name, aes(x = x_var_name, y = y_var_name)) +
  geom_type()
Breaking that down:

First, tell R you’re using ggplot()
Then, tell it the object name where variables exist (data = df_name)
Next, tell it the aesthetics aes() to specify which variables you want to plot
Then add a layer for the type of geom (graph type) with geom_*() - for example, geom_point() is a scatterplot, geom_line() is a line graph, geom_col() is a column graph, etc.
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

```
```{r}
## store base as an object
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

```
```{r}
gg_base +
  geom_point()
```
```{r}
gg_base +
  geom_line()
```
```{r}
gg_base +
  geom_col()
```

```{r}
gg_base +
  geom_area()
```
Some common arguments we’ll use first are:

color = or colour =: update point or line colors
fill =: update fill color for objects with areas
linetype =: update the line type (dashed, long dash, etc.)
pch =: update the point style
size =: update the element size (e.g. of points or line thickness)
alpha =: update element opacity (1 = opaque, 0 = transparent)

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )

```

```{r}
gg_base + 
  geom_point(color = "purple",
             pch = 17,
             size = 4,
             alpha = 0.5)
```
```{r}
gg_base +
  geom_area(aes(fill='orange'))
```

```{r}
gg_base + 
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  )
```

```{r}
gg_base + 
  geom_col(aes(fill = visitors))
```
```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
```
```{r}
#library(plyr)
library(scales)

# ---------------------------------------------------------------------------------------------
# Formatting functions for ggplot  graph axis
# ---------------------------------------------------------------------------------------------

#' Human Numbers: Format numbers so they're legible for humans
#' Use this in ggplot for labels where you might use the comma or percent functions from the 
#' Scales package.
#' 
#' Checks whether numbers are positive or negative. 
#' Allows up to 1 significant figure
#' sapply used for element-wise application of the humanity function as a vector may include
#' numbers where billions, millions or thousands are appropriate.
#'
#' @return a character vector the same length as the input vector
#' @param x a numeric vector to format, 
#' @smbl a symbol you'd like to prefix your numbers by e.g. "$"
#' @signif = the number of significant places you want the function to return
#' @examples
#' human_numbers(c(1000000 , 1500000, 10000000000))
#' human_numbers(c(1.200000e+05, -2.154660e+05, 2.387790e+05, 4.343500e+04 ,5.648675e+12), "$")
#' ggplot2 + scale_y_continuous(labels = human_numbers)
#' ggplot2 + scale_x_continuous(labels = human_numbers)
#' ggplot2 + scale_x_continuous(labels = human_gbp)

human_numbers <- function(x = NULL, smbl ="", signif = 1){
  humanity <- function(y){

    if (!is.na(y)){
        tn <- round(abs(y) / 1e12, signif)
       b <- round(abs(y) / 1e9, signif)
       m <- round(abs(y) / 1e6, signif)
       k <- round(abs(y) / 1e3, signif)

      if ( y >= 0 ){
        y_is_positive <- ""
      } else {
        y_is_positive <- "-"
      }

      if ( k < 1 ) {
        paste0( y_is_positive, smbl, round(abs(y), signif ))
        } else if ( m < 1){
        paste0 (y_is_positive, smbl,  k , "k")
      } else if (b < 1){
        paste0 (y_is_positive, smbl, m ,"m")
      }else if(tn < 1){
          paste0 (y_is_positive, smbl, b ,"bn")
      } else {
        paste0 (y_is_positive, smbl,  comma(tn), "tn")
      }
    } else if (is.na(y) | is.null(y)){
        "-"
    }
  }

  sapply(x,humanity)
}

#' Human versions of large currency numbers - extensible via smbl

human_gbp   <- function(x){human_numbers(x, smbl = "£")}
human_usd   <- function(x){human_numbers(x, smbl = "$")}
human_euro  <- function(x){human_numbers(x, smbl = "€")} 
human_num   <- function(x){human_numbers(x, smbl = "")} 
## https://github.com/fdryan/R/blob/master/ggplot2_formatter.r
```

```{r}
library(scales)
gg_base +
  geom_line(linetype = "dotted") +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)" 
  ) +
  scale_x_continuous(labels = human_num)
```
scale format http://www.cookbook-r.com/Graphs/Axes_(ggplot2)/

combine

```{r}
gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size = year),
             alpha = 0.5) +
  scale_y_continuous(name="Visitors", labels = scales::comma)
+
  theme_minimal()
```
```{r}
gg_base +
  geom_col(fill = "orange",
           color = "purple") +
  geom_line(color = "green")
```
```{r}
ggplot(data = ca_np, aes(x = year, y = visitors)) +
  geom_line()
```
```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line()
```
```{r}
gg_np <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name))
```

```{r}
gg_np +
  geom_line(show.legend = FALSE) +
  theme_light() + 
  labs(x = "year", y = "annual visitors") +
  facet_wrap(~ park_name) + 
  scale_y_continuous(name="Annual Visitors", labels = scales::comma)

```
```{r}
ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 8, height = 7)

```

