---
title: "R for Excel Practice"
output: html_notebook
---
## Import basic library for data cleaning
here: library for pointing files at current project working directory
```{r}
library(tidyverse)
library(readxl)
library(here)
```
```{r}
ca_np <- read_csv(here("data", "ca_np.csv"))
```
```{r}
View(ca_np)
```
```{r}
names(ca_np)
```
```{r}
head(ca_np)
```
```{r}
summary(ca_np)
```
```{r}
library(readxl)
```
```{r}
##ci_np : national parks for Channel Islands
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```
```{r}
View(ci_np)
```

ggplot(data = df_name, aes(x = x_var_name, y = y_var_name)) +
  geom_type()
Breaking that down:

First, tell R you’re using ggplot()
Then, tell it the object name where variables exist (data = df_name)
Next, tell it the aesthetics aes() to specify which variables you want to plot
Then add a layer for the type of geom (graph type) with geom_*() - for example, geom_point() is a scatterplot, geom_line() is a line graph, geom_col() is a column graph, etc.
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

```
```{r}
## store base as an object
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

```
```{r}
gg_base +
  geom_point()
```
```{r}
gg_base +
  geom_line()
```
```{r}
gg_base +
  geom_col()
```

```{r}
gg_base +
  geom_area()
```
Some common arguments we’ll use first are:

color = or colour =: update point or line colors
fill =: update fill color for objects with areas
linetype =: update the line type (dashed, long dash, etc.)
pch =: update the point style
size =: update the element size (e.g. of points or line thickness)
alpha =: update element opacity (1 = opaque, 0 = transparent)

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )

```

```{r}
gg_base + 
  geom_point(color = "purple",
             pch = 17,
             size = 4,
             alpha = 0.5)
```
```{r}
gg_base +
  geom_area(aes(fill='orange'))
```

```{r}
gg_base + 
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  )
```

```{r}
gg_base + 
  geom_col(aes(fill = visitors))
```
```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
```
```{r}
gg_base +
  geom_line(linetype = "dotted") +
  theme_bw() +
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  )
```
scale format http://www.cookbook-r.com/Graphs/Axes_(ggplot2)/

combine

```{r}
gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size = year),
             alpha = 0.5) +
  theme_minimal()
```
```{r}
gg_base +
  geom_col(fill = "orange",
           color = "purple") +
  geom_line(color = "green")
```
```{r}
ggplot(data = ca_np, aes(x = year, y = visitors)) +
  geom_line()
```
```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line()
```
```{r}
gg_np <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name))
```

```{r}
gg_np +
  geom_line(show.legend = FALSE) +
  theme_light() + 
  labs(x = "year", y = "annual visitors") +
  facet_wrap(~ park_name)
```

